---
main_topsize: 0.2 #percent coverage of the poster
main_bottomsize: 0.1
#ESSENTIALS
title: "<span style='font-size: 90px;'>Impact of Country of Origin on  Asylum Case Rejection Rates</span>"
author:
  - name: '**Emory Bayer**'
    affil: 1
    main: true
    orcid: '0000-0002-1099-3857'
    email: eb7666a@american.edu
affiliation:
  - num: 1
    address: School of International Service, American University
output: 
  posterdown::posterdown_html:
    self_contained: TRUE
knit: pagedown::chrome_print
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      warning = FALSE,
                      tidy = FALSE,
                      message = FALSE,
                      fig.align = 'center',
                      out.width = "100%")
options(knitr.table.format = "html") 
```

**Introduction**

As a refugee resettlement caseworker, my interest in this UNHCR dataset is rooted in my professional experience. I worked with clients from Afghanistan and Ukraine who arrived in the United States through temporary immigration pathways, prompting many to seek asylum. This sparked my curiosity about whether there are discernible patterns regarding the denial of asylum applications among individuals from specific countries of origin.

- Anticipated that Ukraine might have higher asylum acceptance rates than Afghanistan; stemmed from belief that racism and Islamophobia still persist in the United States
- Expected applicants from Venezuela, Haiti, and Cuba to have higher asylum rejection rates compared to Afghanistan and Ukraine
- Expectation based on the nature of their asylum claims,less media attention paid to their situations, & historically less government support in creating permanent residency pathways for migrants from these nations

**Expectations:**

- Anticipated that Ukraine might have higher asylum acceptance rates than Afghanistan; stemmed from belief that racism and Islamophobia still persist in the United States
- Expected applicants from Venezuela, Haiti, and Cuba to have higher asylum rejection rates compared to Afghanistan and Ukraine
- Expectation based on the nature of their asylum claims,less media attention paid to their situations, & historically less government support in creating permanent residency pathways for migrants from these nations


 **Importance:**
It is crucial to comprehend this issue within the context of global migration management and to uncover any significant disparities between groups granted asylum and those who are not. Ensuring the fairness and equity of the asylum case system is important, especially considering that individuals applying for asylum are already in vulnerable situations. This makes it even more important that the system remains unbiased and does not discriminate against individuals from specific countries.


Furthermore, it is important to understand this issue from both a legal aid and social service perspective. Analysis of this issue might indicate:

- need for more targeted outreach to individuals from specific countries of origin to assist with their asylum applications

- signal need for reforms within the United States asylum system

- signal need for increased lobbying efforts in Congress to establish legal pathways to permanent residency for citizens from these countries



```{r, include=FALSE}
knitr::write_bib(c('posterdown', 'rmarkdown','pagedown'), 'packages.bib')
```

**Method/Data:**

The data is sourced from the United Nations High Commissioner on Refugees and encompasses asylum case application data from 2015 to 2023. However, for the purpose of this analysis, I chose to filter the data to include only the years 2021 to 2023. It includes several variables which I utilize in this analysis, such as 'Year', 'Country of Origin', 'Country of Asylum', 'Rejected Decisions', and 'Total Decisions'.

*For the purposes of this analysis I created a new variable, entitled "Rejected.percent.overall", to represent the percentage of cases for asylum applicants from each country of origin that were rejected.*

**Rejected.percent.overall** = This variable was created by adding 'Rejected Decisions' and 'Otherwise Closed Decisions' together, then dividing by the number of 'Total Decisions', and finally multiplying by 100. It represents the percentage of asylum case rejections for each 'Country of Asylum' and 'Country of Origin' pair.


I chose to filter the data to only include asylum applicants from Afghanistan, Cuba, Haiti, Ukraine, and Venezuela. Afghanistan and Ukraine were selected due to the significant displacement resulting from ongoing conflicts in those regions. Additionally, I included Haiti, Venezuela, and Cuba because these nationalities have constituted the bulk of asylum applications in the United States over the past two years.

```{r}

library(tidyverse)
library(knitr)
library(patchwork)
library(cowplot)

setwd("C:/Users/emory/Documents/Intro to Data and Statistics/Datasets")
df = read.csv("UNHCR Asylum Data.csv")

World_data <- read.csv("UNHCR Asylum Data.csv")%>%
  filter(Country.of.origin == "Ukraine"| Country.of.origin == "Cuba"| Country.of.origin
         == "Afghanistan"| Country.of.origin == "Haiti"| Country.of.origin ==
           "Venezuela (Bolivarian Republic of)")%>%
  filter(Cases...Persons == "P")%>%
  filter(Year>= 2021)%>%
  group_by(Country.of.asylum, Country.of.origin, Year)%>%
  summarize(Recognized.decisions = sum(Recognized.decisions), 
            Complementary.protection = sum(Complementary.protection),
            Rejected.decisions = sum(Rejected.decisions),
            Otherwise.closed = sum(Otherwise.closed), 
            Total.decisions = sum(Total.decisions))%>%
  mutate(
    Rejected.percent.overall = ((Rejected.decisions + Otherwise.closed) / Total.decisions) * 100,
    Rejected.case.percent = (Rejected.decisions / Total.decisions) * 100,
    Rejected.admin.percent = (Otherwise.closed / Total.decisions) * 100)


asylum_table <- World_data %>%
  group_by(Country.of.origin, Year) %>%
  summarize(Total_Asylum_Seekers = sum(Total.decisions))



asylum_table_rejections <- World_data%>%
  group_by(Country.of.origin, Year) %>%
  summarise(Total_asylum_case_rejections = sum(Rejected.decisions))



 country_colors <- c(
    "Afghanistan" = "#1f77b4", 
    "Ukraine" = "#ff7f0e", 
    "Haiti" = "#2ca02c", 
    "Venezuela (Bolivarian Republic of)" = "#d62728",
    "Cuba" = "#9467bd"
  )
  
  asylum_table$Country_Year <- paste(asylum_table$Country.of.origin, asylum_table$Year, sep = "_")
  
  
  asylum_table_rejections$Country_Year <- paste(asylum_table_rejections$Country.of.origin, asylum_table_rejections$Year, sep = "_")
  
  
  desired_order <- c(
    "Afghanistan_2021", "Afghanistan_2022", "Afghanistan_2023",
    "Cuba_2021", "Cuba_2022", "Cuba_2023",
    "Haiti_2021", "Haiti_2022", "Haiti_2023",
    "Ukraine_2021", "Ukraine_2022", "Ukraine_2023",
    "Venezuela (Bolivarian Republic of)_2021", "Venezuela (Bolivarian Republic of)_2022", "Venezuela (Bolivarian Republic of)_2023"
  )
  
  asylum_table$Country_Year <- factor(asylum_table$Country_Year, levels = desired_order)
  asylum_table_rejections$Country_Year <- factor(asylum_table_rejections$Country_Year, levels = desired_order)
  
  
  plot_total_asylum <- ggplot(asylum_table, aes(x = Country_Year, y = Total_Asylum_Seekers, fill = Country.of.origin)) +
    geom_bar(stat = "identity", position = "dodge", colour = "black") +
    geom_text(aes(label = Year), position = position_dodge(width = 1), vjust = 1.5, size = 3) +  
    labs(x = "Year", 
         y = "Asylum Applications",
         fill = "Country of Origin") +
    scale_fill_manual(values = country_colors) +
    theme_minimal() +
    theme(axis.text.x = element_blank(),  
          axis.title.x = element_text(size = 12),  
          axis.text.y = element_text(size = 8),
          panel.border = element_rect(color = "black", fill = NA, size = 1),
          plot.title = element_text(size = 12)) +  
    coord_cartesian(clip = "off") +  
    scale_x_discrete(expand = c(0.1, 0))  
  
  plot_rejections <- ggplot(asylum_table_rejections, aes(x = Country_Year, y = Total_asylum_case_rejections, fill = Country.of.origin)) +
    geom_bar(stat = "identity", position = "dodge", colour = "black") +
    geom_text(aes(label = Year), position = position_dodge(width = 1), vjust = 1.5, size = 3) +  
    labs(x = "Year",  
         y = "Asylum Case Rejections") +
    scale_fill_manual(values = country_colors) +
    theme_minimal() +
    theme(axis.text.x = element_blank(),  
          axis.title.x = element_text(size = 12),  
          axis.text.y = element_text(size = 8),
          panel.border = element_rect(color = "black", fill = NA, size = 1),
          plot.title = element_text(size = 12)) +  
    theme(legend.position = "none") +
    coord_cartesian(clip = "off") +  
    scale_x_discrete(expand = c(0.1, 0))  
  
  combined_plots <- plot_total_asylum / plot_rejections
  
  
  combined_plots_with_title <- ggdraw() +
    draw_label("Comparison of Total Asylum Applications and Case Rejections", 
               x = 0.5, y = 0.98, 
               hjust = 0.5, vjust = 1, 
               size = 14,  
               fontface = "bold")
  
  
  final_plot <- cowplot::plot_grid(
    combined_plots_with_title,
    combined_plots,
    ncol = 1,
    rel_heights = c(0.05, 1)  
  )
  
  
  print(final_plot)
```
 **Results**

In terms of the results of the regression, the base 'Country of Origin' which each other 'Country of Origin' is being compared to is Afghanistan. The results indicate that for every other 'Country of Origin', they have statistically significant higher asylum case rejection rates than those from Afghanistan. This can be observed in the visual below.

The visual depicts the asylum rejection rates by 'Country of Origin' using the parameters of the regression. The black line running through each box indicates the median asylum rejection percentage, with the bottom of the box indicating the 1st quartile, or 25% of the data, and the top of the box indicating the 3rd quartile of the data, or 75%.


This visual indicates that: 

- Asylum seekers from Cuba are the most likely to have their asylum cases rejected, with a median asylum rejection rate of~77%
- Asylum applicants from Haiti have a median asylum rejection rate of 71%
- Asylum applicants from Venezuela have a median asylum rejection rate of ~ 60%
- Asylum applicants from Ukrainian have a median asylum rejection rate of ~58%
- Asylum applicants from Afghanistan have a median asylum rejection rate of ~40% 
```{r}
library(tidyverse)
library(knitr)
library(kableExtra)

setwd("C:/Users/emory/Documents/Intro to Data and Statistics/Datasets")
df = read.csv("UNHCR Asylum Data.csv")

World_data <- read.csv("UNHCR Asylum Data.csv")%>%
  filter(Country.of.origin == "Ukraine"| Country.of.origin == "Cuba"| Country.of.origin
         == "Afghanistan"| Country.of.origin == "Haiti"| Country.of.origin ==
           "Venezuela (Bolivarian Republic of)")%>%
  filter(Cases...Persons == "P")%>%
  filter(Year>= 2021)%>%
  group_by(Country.of.asylum, Country.of.origin, Year)%>%
  summarize(Recognized.decisions = sum(Recognized.decisions), 
            Complementary.protection = sum(Complementary.protection),
            Rejected.decisions = sum(Rejected.decisions),
            Otherwise.closed = sum(Otherwise.closed), 
            Total.decisions = sum(Total.decisions))%>%
  mutate(
    Rejected.percent.overall = ((Rejected.decisions + Otherwise.closed) / Total.decisions) * 100,
    Rejected.case.percent = (Rejected.decisions / Total.decisions) * 100,
    Rejected.admin.percent = (Otherwise.closed / Total.decisions) * 100)


Broad_Regression<-lm(data = World_data, formula = Rejected.percent.overall ~ Country.of.origin+as.factor(Year))

predictions <- data.frame(World_data$Country.of.origin, predict(Broad_Regression))


colnames(predictions) <- c("Country.of.origin", "Predicted_Rejection")


country_colors <- c("Afghanistan" = "#1f77b4", 
                    "Ukraine" = "#ff7f0e", 
                    "Haiti" = "#2ca02c", 
                    "Venezuela (Bolivarian Republic of)" = "#d62728",
                    "Cuba" = "#9467bd")


boxplot_plot <- ggplot(predictions, aes(x = Country.of.origin, y = Predicted_Rejection, fill = Country.of.origin)) +
  geom_boxplot() +
  labs(x = NULL, y = "Predicted Rejection Percent", title = "Visualization of Regression Results: 
       Impact of Country of Origin on 
       Asylum Case Rejection Percentages")+
  scale_fill_manual(values = country_colors, name = "Country of Origin") + 
  theme_minimal()+
  theme(axis.text.x = element_blank(),  
        panel.border = element_rect(color = "black", fill = NA, size = 1),
        plot.title = element_text(face = "bold"))


boxplot_plot

```

**Findings**
**Findings**

The regression has shown a statistically significant relationship between the country of origin of asylum applicants and the percentage of asylum cases which are rejected. Specifically, Afghan asylum seekers have significantly lower asylum rejection rates than applicants from other countries of origin. This is exemplified by the p-value for every country of origin aside from Afghanistan being below 0.001. P-values by country were:    

- Cuba: < 2e-16
- Haiti: < 2e-16
- Ukraine: 6.01e-06
- Venezuela: 3.91e-06

This indicates that the chances of observing these changes in asylum rejection rates, if there was no relationship between the country of origin being Afghanistan and a different country of origin, would be less than 0.001.

This is interesting to me and makes me curious about what factors make applications for Afghan asylees more likely to have their applications approved. It could be purely based on the legal grounds for their asylum applications. However, other factors, such as their ability to access adequate legal services or whether they are eligible for deportation under a "safe third-party agreement," could influence why asylees from some countries are much more likely to have their applications denied.

```{r, iristable}

```

# References

UNHCR. (n.d.). Refugee Data Finder. UNHCR. https://www.unhcr.org/refugee-statistics/download/
?url=5lq2In 

Roy, D. (n.d.). Seeking protection: How the U.S. asylum 
process works. Council on Foreign Relations. https://tinyurl.com/mry6t6a8

